Sequence Specific Modeling of \emph{E.~coli} Cell-Free Protein Synthesis

### 2017-05-05 Paper_v1.tex
Small changes

In this study, we used sequence specific constraint based modeling to evaluate the performance of \emph{E.~coli} cell-free protein synthesis (CFPS).
A core \emph{E.~coli} cell-free metabolic model was developed from literature \cite{Feist:2007aa}.
This model, which described glycolysis, pentose phosphate pathway, amino acid biosynthesis and degradation and energy metabolism, was then augmented with
sequence specific descriptions of promoter function, transcription and translation processes.
Thus, sequence specific constraints based modeling explicitly coupled transcription and translation with the availability of metabolic resources.
We tested this approach by simulating the production of two model proteins, and then investigated the productivity, energy efficiency, and carbon yield for eight different proteins.
From this, higher carbon number proteins typically had lower productivity rates and carbon yields than that of the lower carbon number proteins; whereas energy efficiency was independent of protein size.
Our model correlations were independently validated with a protein not in our original data set.
We compared the theoretical optimum solutions of ssFBA with an experimentally constrained case which suggested CFPS retained a growth-like operational memory that led to over consumption of glucose, and over production of ATP.
Further, global sensitivity analysis identified the key metabolic processes that controlled circuit performance, showing oxidative phosphorylation as instrumental for maintaining a high energy efficiency and carbon yield and the translation rate for productivity.
Taken together, sequence specific constraints based modeling offers a novel means to \emph{a~priori} estimate the performance of cell-free synthetic circuits.

@@

The carbon yield for experimentally constrained CAT production was 6.2\% compared to the theoretical maximum of 58\% (Fig.~\ref{fig:Prof}F, dagger).
Given that the CAT productivity was similar between the simulated and measured systems, differences in
the glucose consumption rate and byproduct accumulation were likely responsible for the difference between the optimal and experimental systems.
The translation rate has been identified as the rate-limiting step for cell-free protein synthesis \cite{2005_underwood_biotech,2014_li_PlosOne}, which was confirmed by the translation rate flux always hitting the upper bound in the ssFBA simulation, regardless of protein.
Since CAT productivity could not be improved, the increased carbon consumption accumulated in byproducts such as pyruvate, acetate, lactate, succinate and malate.
In addition, the accumulation of acetate and lactate suggested that oxidative phosphorylation activity was limited.
Thus to improve carbon yield for CFPS, the translation rate should be improved for increased protein productivity, or energy efficiency should be improved to lower the glucose consumption rate.
%%%%%%%%%%%%%%%%%%%%
%However, it may also be hypothesized that only the amino acids supplied contribute to the carbon yield of the protein of interest, while glucose is used to energize CFPS.
%If this is the case, the carbon yield for the optimal and experimental systems resulted in approximately 100\%.
%%%%%%%%%%%%%%%%%%%%%

>>>>

In this study, we used sequence specific constraint based modeling to evaluate the performance of \emph{E.~coli} cell-free protein synthesis (CFPS).
A core \emph{E.~coli} cell-free metabolic model was developed from literature \cite{Feist:2007aa}.
This model, which described glycolysis, pentose phosphate pathway, amino acid biosynthesis and degradation and energy metabolism, was then augmented with
sequence specific descriptions of promoter function, transcription and translation processes.
Thus, sequence specific constraints based modeling explicitly coupled transcription and translation with the availability of metabolic resources.
We tested this approach by simulating the production of two model proteins, and then investigated the productivity, energy efficiency, and carbon yield for eight different proteins.
From this, higher carbon number proteins typically had lower productivity rates and carbon yields than that of the lower carbon number proteins; whereas energy efficiency was independent of protein size.
Our model correlations were independently validated with a protein not in our original data set.
We compared the theoretical optimal solutions with an experimentally constrained case which suggested CFPS retained a growth-like operational memory that led to over consumption of glucose, and over production of ATP.
Further, global sensitivity analysis identified the key metabolic processes that controlled circuit performance, showing oxidative phosphorylation as instrumental for maintaining a high energy efficiency and carbon yield and the translation rate for productivity.
Taken together, sequence specific constraints based modeling offers a novel means to \emph{a~priori} estimate the performance of cell-free synthetic circuits.

@@

The carbon yield for experimentally constrained CAT production was 6.2\% compared to the theoretical maximum of 58\% (Fig.~\ref{fig:Prof}F, dagger).
Given that the CAT productivity was similar between the simulated and measured systems, differences in
the glucose consumption rate and byproduct accumulation were likely responsible for the difference between the optimal and experimental systems.
The translation rate has been identified as the rate-limiting step for cell-free protein synthesis \cite{2005_underwood_biotech,2014_li_PlosOne}, which was confirmed by the translation rate flux always hitting the upper bound in the simulation, regardless of protein.
Since CAT productivity could not be improved, the increased carbon consumption accumulated in byproducts such as pyruvate, acetate, lactate, succinate and malate.
In addition, the accumulation of acetate and lactate suggested that oxidative phosphorylation activity was limited.
Thus to improve carbon yield for CFPS, the translation rate should be improved for increased protein productivity, or energy efficiency should be improved to lower the glucose consumption rate.
%%%%%%%%%%%%%%%%%%%%
%However, it may also be hypothesized that only the amino acids supplied contribute to the carbon yield of the protein of interest, while glucose is used to energize CFPS.
%If this is the case, the carbon yield for the optimal and experimental systems resulted in approximately 100\%.
%%%%%%%%%%%%%%%%%%%%%
